### 2017-09-28 article.tex
primitives
recovering read and write

\documentclass{article}

\usepackage{amsmath}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}

%opening
\title{ISF Grant}
\author{}

\begin{document}

\maketitle

\begin{abstract}
	a
\end{abstract}

\section{Introduction}

Traditional computer architectures containing both a volatile main memory alongside a non-volatile secondary storage, have led to a separation of program state into operational data stored using in-memory data structures, and recovery data stored using sequential on-disk structures such as transaction logs. A system-wide failure, such as a power outage, cause a full lost of the in-memory data structures, which must be reconstructed entirely from the recovery data, making the software system temporarily unavailable. As a result, the design of in-memory data structures emphasise disposable constructs optimized for parallelism, in contrast to the structures that hold recovery data, which cannot be discarded upon a failure and which benefit less from parallelism since their performance is limited by the secondary storage.

Current industry projections indicate that non-volatile main memory (NVRAM) will become commonplace over the next few years. While the availability of NVRAM suggests the possibility of keeping persistent data in main memory (not just in the file system), the fact that recent updates to registers and cache may be lost during a power failure means that the data in main memory, if not carefully managed, may not be consistent at recovery time.
Traditional log-based recovery techniques, although can be applied correctly in such systems, fail to take full advantage of the parallelism enabled by allowing processing cores to access recovery data directly using memory operations rather than slow block transfers from secondary storage. As a result, harnessing the performance benefits of NVRAM-based platforms requires a careful rethinking of recovery mechanisms.

The study of shared data structures and concurrent programming dates back over a half century ago. 1965 Dijkstra's pioneering work presented the concept of concurrent programming [*], followed by a series of seminal papers on inter-process communication, wait-free synchronization, and linearizability [*].
Recent research has paid close attention to asynchronous models, in which there is no bound on the amount of time a process takes to transition to its next step, or to complete a memory operation. This assumption captures the behaviour of most modern memory hierarchies, where different media (e.g., L1 cache vs. L2 cache vs. main memory) incur vastly different and often unpredictable access latencies, as well as the effect of the operating system (e.g., via preemption and interrupts) on the liveness of processes. In addition, asynchrony is also related to reliability in a precise way: algorithms that provide non-blocking progress properties (e.g., lock-freedom and wait-freedom) in an asynchronous environment with reliable processes continue to provide the same progress properties if crash failures are introduced. Informally speaking, this property holds because a process that crashes permanently at an arbitrary point in the execution of its algorithm is indistinguishable to the other processes from one that is merely very slow.

\subsection{Correctness properties}
Correctness property specifies the requirements for an implementation to be considered correct. Herlihy and Wing’s linearizability property [*] is widely adopted for shared objects in conventional shared memory models.  Roughly speaking, linearizability requires that operations on objects appear to take effect instantaneously in some serial order, while preserving the happens before relation, i.e., if $op_1$ ends before $op_2$ begins, then $op_1$ should precede $op_2$ in the serial order. Linearizability has no notion of an aborted or failed operation, and requires that a process finish one operation before it invokes the next. The criteria according to which implementations that support crash-recovery failures are correct remains unclear.

Aguilera and Frølund have proposed strict linearizability as a safety condition for persistent concurrent objects, which treats the crash of a process as a response, either successful or unsuccessful, to the interrupted operation [*]. A successful response means that the operation takes effect at some point between its invocation and the crash failure, and an unsuccessful response means that the operation does not take effect at all. Strict linearizability preserves both locality and program order. However, they proved that it precludes the implementation of some wait-free objects for certain machine models - there is no wait-free implementations of multi-reader single-writer (MRSW) registers from single-reader single-writer (SRSW) registers under strict linearizability.

Guerraoui and Levy have proposed persistent atomicity [*]. It is similar to strict linearizability, but allows an operation interrupted by a failure to take effect before the subsequent invocation of the same process, possibly after the failure. They also proposed transient atomicity, which relaxes this criterion even further and allows an interrupted operation to take effect before the subsequent write response of the same process. Both conditions ensure that the state of an object will be consistent in the wake of a crash, but it does not provide locality: correct histories of separate objects, when merged, will not necessarily yield a correct composite history.

Berryhill et al. have proposed an alternative, recoverable linearizability [*], which achieves locality but may sacrifice program order after a crash. It is a relaxed version of persistent atomicity, which requires the operation to be linearizes or aborts before any subsequent linearization by the pending thread on that same object.

Izraelevitz et al. considered a real-world failure model, in which processes are assumed to fail together, as part of a full-system crash. Under this model, persistent atomicity and recoverable linearizability are indistinguishable (and thus local). The term durable linearizability was used to refer to this merged safety condition under the restricted failure model.

\subsubsection{Recoverable response}
All the correctness properties discussed above focused on keeping the object or the data structure consist in case of a failures. However, none of which address the problem of the returned value by the failed operation. Consider for example a write operation. Since write is atomic it obviously satisfies strict linearizability - either the write took effect before the crash failure, or that it did not. However, once the process is recovered it has no knowledge which of the two is the case, since the acknowledgement returned by the write is saved to the local memory of the process (a part of the cache), which is lost due to the failure. Assuming the process is in a middle of executing an operation which it desired to complete once it is recovered, it must know whether the write took effect (in which case it can continue and execute the code), or that it did not (in which case it needs to rewrite). Skipping the write is not an option, as it might cause an undesired behaviour, which in turn might violate the requirements of the object.

For this case, we present a different approach called recoverable response linearizability (RR-linearizability) (good term????). In case of a failure, once the process is recovered it is "aware" of the failure, in the sense that it runs a designated recovery code. In a sense, this recovery code "extends" the interval of the interrupted operation. We require that the pending operation linearizes between it's invocation and the end of the recovery code, or aborts. In addition, in case of linearization, the returned value of the operation is known to the process. Notice that this definition introduce a relaxation which is not intuitive neither trivial - in case that a process operation is complete before the crash, but the response of it is lost due to the cache's lost, we allow the process, upon recovering, extending the interval of that operation using the recovery code (even though the operation was complete), such that at the end of it the process knows the returned value or aborts. This gives the process the flexibility to move the linearization point to after the crash, if needed. In addition, although the definition allows a process to abort the interrupted operation, it is equivalent to a definition in which a process must complete the operation - once a process knows the operation is aborted, it can reissue it.

RR-linearizability does not specify the progress property required from the implementation. One can always use a recoverable mutual exclusion lock at the recovery code in order to sequentially let the processes complete their pending operations. Obviously, this solution is not desired when the operation implementation we are recovering is wait-free. In this case, we would like the recovery code to be wait-free as well, meaning that a process finish it in a finite number of its own steps in the absence of failure. One more point that requires our attention is the possibility of a failure along the recovery code. Would it be realistic to assume that such a scenario will not happen? Of course, a recovery code that can suffer itself failures is better. However, this is not well defined. If we allow each failure to run new and different recovery code, we can use a nesting technique in a way which increase the running time with each failure. A better approach will be such that in the case of failure inside of a recovery code, restarting the recovery code, from the beginning any other point, is enough. In such case, if the recovery code is wait-free in the absence of a failure, we can guarantee that no matter how many failures the process experienced, once given enough time with no failure, it will finish the recovery code (and therefore the pending operation) in a wait-free manner.

\subsubsection*{Recoverable primitives}

The correctness properties variants discussed earlier do not need to consider primitives, as any atomic operation satisfies any of the properties. Unlike it, RR-linearizability is not satisfies by primitives since the respond of the primitive is lost in case of a failure. Therefore, there is a need to supply each primitive with an implementation as well as a recovery code. In the following section we do so for well known primitives, as well as proving impossibility result for others. We assume a failure model of a full-system crash, and address the wait-free progress property.
The fact that RR-linearizabilty allows us to swift the linearization point of an operation to after the crash is used for the recover after a primitive failure.

\paragraph*{Read}

The process will simply reread the variable upon its recovering. Since read does not affect the rest of the processes, this can be seen as aborting the read preceding the crash, and reissue it.

\paragraph*{Write}
Write instruction is "wrap" with code such that in case of a failure the extra data will be used for recovering. For an instruction writing value $x$ to variable $R$ by process $p$, we provide the following implementation. We use the convention of capital letters names for shared memory variables, and small letters for local ones. In the following code, $R_p$ is a variable in the memory designated for process $p$.

\begin{algorithm}
	\caption{Write}\label{euclid}
	\begin{algorithmic}[1]
		\Procedure{write operation}{}
		\State $res \gets R$
		\State $R_{p} \gets res$
		\State $R \gets x$
		\EndProcedure

		\Procedure{recovery code}{}
		\If {$R_{p} == R$} \Return abort
		\EndIf
		\EndProcedure
	\end{algorithmic}
\end{algorithm}

For simplicity, we write the recovery code as a single instruction, although it needs to be written as several instructions, as it access two different locations at the memory. Since $R_p$ is accessed by $p$ only, the point where $p$ reads $R$ determine the outcome of the recovery code. Moreover, in case of a failure at the recovery code, upon recovering the process will execute the recovery code again.

In case of a failure before line 4, the process will abort upon recovering (and reissue the write operation). In case of a failure at line 4, the process will execute the recovery code. Assuming the if statement at the recovery code does hold, then no process, including $p$, wrote to $R$ after the read of $p$ at line 2, and thus the write operation is aborted and needs to be reissued. Otherwise, there must be a process to write to $R$ after the read at line 2. In case $p$ write at line 4 took affect before the crash  we linearize it at the point of the write. Otherwise there must be a process $q \neq p$ such that $p$ wrote to $R$ in between the two reads of $p$, and we linearize the write of $p$ just before $q$'s write. Notice that the write of $q$ covers the write of $p$, and therefore other processes can not distinguish the new execution with the linearized write from the run with no write by $p$.

The above analysis ignores the ABA problem - it might be that $p$ reads the value from $R$, even though there was a write to $R$ in between the two different reads. To overcome this problem, we can augment any value written with the writing process's id, and a sequential number (each process will have its own sequential number). This way, reading the same value guarantee that no write to $R$ took place between the two different reads.

\bibliographystyle{abbrv}
\bibliography{references}

\end{document}